---
title: "p8105_hw2_yh3555"
author: "Yuchen Hua"
date: "2022-10-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(haven)
library(dplyr)
library(readxl)
library(tidyr)
```

## Problem 1

Read the dataset and retain the variables. 

```{r}
nyc_trans =
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
    janitor::clean_names() %>%
    select(
      line, station_name, station_latitude, station_longitude, starts_with("routes"), 
      entry, exit_only, vending, entrance_type, ada) %>%
    mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

To find distinct stations. 
```{r}
nyc_trans %>%
  select(station_name, line) %>%
  distinct
```

To find ADA-compliant stations.
```{r}
nyc_trans %>%
  filter(ada == TRUE ) %>%
  select(station_name, line) %>%
  distinct
```

To compuate entrance/ exist without vending allowing entrance. 
```{r}
nyc_trans %>%
  filter(vending == "NO") %>%
  pull(entry) %>%
  mean
```

## Problem 2
To take a look at the the sheets in the excel file. 
```{r}
excel_sheets("./data/Trash Wheel Collection Data.xlsx")
```

Manipulate Mr. Trash wheel sheet.  
```{r}
mr.trash = read_excel("./data/Trash Wheel Collection Data.xlsx", 
                        sheet = "Mr. Trash Wheel", 
                        range = "A2:N549") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))

mr.trash
```

Manipulate Professor Trash Wheel sheet. 
```{r}
pro.trash = read_excel("./data/Trash Wheel Collection Data.xlsx", 
                       sheet = "Professor Trash Wheel",
                       range = "A2:M96") %>%
  janitor::clean_names() %>%
  drop_na(dumpster)

pro.trash
```

To add variables to both dataset for combination.  

```{r}
mr.trash = mutate(mr.trash, label = "m.t") %>%
  mutate( year = as.double(year))

pro.trash = mutate (pro.trash, label = "p.t")

total_trash = full_join(mr.trash, pro.trash)

total_trash
```

## Problem 3


Import and clean ip data of pols-month.csv. 
```{r}
pols_month = 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col= mon, into =  c('year', 'month', 'day'))%>%
  mutate( 
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
    day = as.integer(day)) %>%
  mutate(president = ifelse(prez_dem == 1, 'dem', 'gop') )%>%
  select(-prez_dem, -prez_gop, -day)

pols_month
```


Clean up the the data in snp.csv
```{r}
snp = 
  read_csv("./data/snp.csv") %>%
  separate(col = date, into = c('month', 'day', 'year')) %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
    day = as.integer(day),
    )%>%
  mutate(year = ifelse( year > 49, year + 1900, year + 2000))%>%
  arrange(year, month) %>%
  select(year, month, close, -day)
         
snp
```



Clean up the Unemployment data
```{r}
Unemp = 
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer( jan : dec,
                names_to = 'month',
                values_to = "Unemployment rate") %>%
  mutate(year = as.integer(year))
  
Unemp
```

Time to merge snp into pols_month. and merge unemployment into the result. 
```{r}
pols_snp = 
  left_join(pols_month, snp, by = c("year", "month"))

pols_snp
```

```{r}
total = 
  left_join(pols_snp, Unemp, by = c("year", "month"))

total
```
